Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 egpaper_final.tex

\section{Related Works}
\subsection{Whitened Histograms of Orientations (WHO)}
% History of Whitening and how whitening works
Recently, Hariharan \etal. introduced Whitened Histograms of Orientations (WHO)
\cite{Hariharan12}, which used statistics from natural images to whiten HOG
templates, making them more discriminative. Whitening is a common signal
processing operation for decorrelating a set of random variables
\cite{Martinsson05, Belouchrani00}. More formally, suppose that we have a
$k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By
whitening the signal, i.e., $\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X])$, we
remove 2nd order correlation between all features. So, all elements in the
random variable $\tilde{X} = \left[\tilde{X}_1, \tilde{X}_2, \dots,
\tilde{X}_k\right]$, become more independent. In the context of HOG features, a
whitened HOG template will produce Gaussian noise when convolved with a random
natural image, but a delta-like response on an image matching the template. WHO
templates are more discriminative than standard HOG templates because they only
capture the gradients that differentiate the object from random images.
\scream{chris added the following} Hariharan \etal. also interpreted the
whitening as Linear Discriminant Analysis (LDA) where all classes have different
means but have the same covariance. The whitening and LDA have the same
formulation when we make a decision boundary $w = \Sigma^{-1}(\mu_+ - \mu_0)$
where $\mu_+$ is the features from a class (LDA) or a signal to whiten
(whitening), $\mu_0$ is the negative class center (LDA) or a signal statistics
(whitening) and $\Sigma$ is the covariance matrix of classes (LDA) or
correlation within signal (whitening).

\textcolor{red}{Should we mention how our method improves on this (ie by using CG)?}

@@

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both more discriminative than WHO, and can be generated in 70ms (compared to several seconds for WHO). This speed up means we can generate templates on the fly, allowing our approach to evaluate arbitrary viewpoint hypotheses. 
\scream{chris:We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), an improvement over WHO which is both more discriminative than WHO and can be computed many orders of magnitude faster than WHO (70ms as opposed to to several seconds for WHO). This speedup allows us to generate NZ-WHO templates on the fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

@@

Our second innovation is 'Non-Zero' whitening. When synthesizing detection templates from rendered images, a common problem is how to handle the background. If we render the model over a natural image background, gradients in the background will be incorporated into the discriminative template. Alternatively, we could leave the background textureless \scream{did Hariharan do this?} (see Fig.~\ref{fig:rendering}), but whitening the resulting HOG template produces in undesirable background artifacts. These are created when centering the template (by subtracting the mean $\mu$), where strong negative weights are introduced in the textureless region (as seen in Fig.~\ref{fig:whocomparison}). This could result in positive matches being suppressed due to spurious background gradients.

@@

Table \ref{tab:who_initializations} provides a detailed comparison of WHO and NZ-WHO templates, illustrating that our method is faster and more discriminative than Whitened Histograms of Orientations. \scream{Does the table show this?}
% We compare the performance of WHO and NZ-WHO on 3D Object dataset \cite{Savarese07} and report performance on Table \ref{tab:who_initializations}.

@@

Figure \ref{fig:spectral} shows the singular values of a covariance matrix for
one HOG template after a small regularization constant is added to the diagonal.
The condition number of of the matrix is \scream{x}, much smaller than the size
of the matrix (7500). As a result, conjugate gradient converges in \scream{x
ms}, two orders of magnitude faster than the system can be solved using the
Cholesky factorization (\scream{5 s}).

The final challenge in on-the-fly template evaluation is performing the
convolution between the template and the target image. Standard convolution
would be prohibitively slow for the high resolution templates we're using, so we
implemented the convolution via Fast Fourier Transform on the GPU. This allows
template evaluation to run in \scream{x ms}, compared to \scream{x ms} for the
naive CPU implementation.

@@

\begin{comment}
\subsection{High Resolution Templates and FFT based Convolution}
\label{sec:fft} 
We generate high resolution templates with more than 250 HOG cells to capture details of an object to give accurate 2D-3D matching. These large templates cause computational burden when computing convolution. Though good for detecting accurate model and pose, convolution of high-resolution templates are much more slower since computation time scales linearly with the number of HOG cells in the template. To overcome and speed up the convolution, we propose FFT based GPU convolution which scales. Suppose there are length $n$ signal and length $m$ filter, naive convolution takes $O(nm)$ time where as FFT-based convolution takes $O\left( (n + m)\log (n+m) \right)$ time. For large $m$, high resolution template, we can gain computational advantage.

we pad zero to the 2D image features by the largest kernel size. And we transform the kernel to match the padded image feature. To compute convolution, we compute element-wise product with the transformed kernel and sum all the features. Unlike \cite{Podlozhnyuk}, we do not need to shift the data nor kernel if we use simple mathematical trick.

\begin{align}
    x_{pad} & = \mathcal{P}_{n+m}(x)\\
    y_{pad} & = \mathcal{P}_{n+m}(y)\\
    x_{pad} \ast y_{pad} & = \mathcal{F}^{-1}(\mathcal{F}(x_{pad}) \circ \mathcal{F}(y_{pad}))
\end{align}

  Where $\ast$ is the circular convolution and $\circ$ is the Hadamard Product and $\mathcal{F}$ is the Fourier Transformation and $\mathcal{P}_n$ is the padding operation that append zeros to make vector of size $n$.

For each dimension of features, we transform kernel and image feature into frequency domain using Fast Fourier Transformation. We transform the image into frequency domain once and use the transformed image for all templates.

We implemented the convolution using FFT on GPU and got 10x speed up.

[Figure for convolution time for different template resolution, CPU, GPU, naive convolution]
\end{comment}
%------------------------------------------------------------------------

\section{Estimating Parameters using MCMC}
\label{sec:fine}
% Estimating viewpoint continuously to fine tune a initial matching
The NZ-WHO template matching method we have presented makes template generation
and evaluation computationally inexpensive. This means we can efficiently
explore the parameter space to find the best object pose, scale, model type and
camera focal length.

Our method begins by convolving the

@@

% there are several advantages that was not possible such as \textbf{on-the-fly} template generation and making large number of templates  
\begin{figure}[t]
\centering
    \includegraphics[width=0.7\linewidth]{tuning2} \\ [-5pt]
    \caption{Different modes of Metropolis Hasting proposals, see Section \ref{sec:fine}}
 \label{fig:tuningmode}
\end{figure}

>>>>

\section{Related Works}
\subsection{Whitened Histograms of Orientations (WHO)}

\scream{Should we merge this with NZ-WHO?}
% History of Whitening and how whitening works
Recently, Hariharan \etal. introduced Whitened Histograms of Orientations (WHO)
\cite{Hariharan12}, which used statistics from natural images to whiten HOG
templates, making them more discriminative. Whitening is a common signal
processing operation for decorrelating a set of random variables
\cite{Martinsson05, Belouchrani00}. More formally, suppose that we have a
$k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By
whitening the signal, i.e., $\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X])$, we
remove 2nd order correlation between all features. So, all elements in the
random variable $\tilde{X} = \left[\tilde{X}_1, \tilde{X}_2, \dots,
\tilde{X}_k\right]$, become more independent. In the context of HOG features, a
whitened HOG template will produce Gaussian noise when convolved with a random
natural image, but a delta-like response on an image matching the template. WHO
templates are more discriminative than standard HOG templates because they only
capture the gradients that differentiate the object from random images.
\scream{requires revision} Hariharan \etal. also interpreted the
whitening as Linear Discriminant Analysis (LDA) where all classes have different
means but have the same covariance. The whitening and LDA have the same
formulation when we make a decision boundary $w = \Sigma^{-1}(\mu_+ - \mu_0)$
where $\mu_+$ is the features from a class (LDA) or a signal to whiten
(whitening), $\mu_0$ is the negative class center (LDA) or a signal statistics
(whitening) and $\Sigma$ is the covariance matrix of classes (LDA) or
correlation within signal (whitening).

\textcolor{red}{Should we mention how our method improves on this (ie by using CG)? Since we explain in depth later, it would be okay}

@@

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both
more discriminative than WHO, and can be generated in 70ms (compared to several
seconds for WHO). This speed up means we can generate templates on the fly,
allowing our approach to evaluate arbitrary viewpoint hypotheses. 
\scream{require revision: We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO),
an improvement over WHO which is both more discriminative than WHO and can be
computed many orders of magnitude faster than WHO (70ms as opposed to to
several seconds for WHO). This speedup allows us to generate NZ-WHO templates
on the fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

@@

Our second innovation is 'Non-Zero' whitening. When synthesizing detection
templates from rendered images, a common problem is how to handle the
background. If we render the model over a natural image background, gradients
in the background will be incorporated into the discriminative template.
Alternatively, we could leave the background textureless \scream{did Hariharan
do this? Nobody did this before} (see Fig.~\ref{fig:rendering}), but whitening the resulting HOG
template produces in undesirable background artifacts. These are created when
centering the template (by subtracting the mean $\mu$), where strong negative
weights are introduced in the textureless region (as seen in
Fig.~\ref{fig:whocomparison}). This could result in positive matches being
suppressed due to spurious background gradients.

@@

Table \ref{tab:who_initializations} provides a detailed comparison of WHO and NZ-WHO templates, illustrating that our method is faster and more discriminative than Whitened Histograms of Orientations. \scream{Does the table show this? should we do more than AP and realtime analysis?}
% We compare the performance of WHO and NZ-WHO on 3D Object dataset \cite{Savarese07} and report performance on Table \ref{tab:who_initializations}.

@@

% Figure \ref{fig:spectral} shows the singular values of a covariance matrix for
% one HOG template after a small regularization constant is added to the diagonal.
% The condition number of of the matrix is , much smaller than the size
% of the matrix (7500). As a result, conjugate gradient converges in 50
% ms, two orders of magnitude faster than the system can be solved using the
% Cholesky factorization (5 sec).

% The final challenge in on-the-fly template evaluation is performing the
% convolution between the template and the target image. Standard convolution
% would be prohibitively slow for the high resolution templates we're using, so we
% implemented the convolution via Fast Fourier Transform on the GPU. 
% 
% This allows
% template evaluation to run in \scream{x ms}, compared to \scream{x ms} for the
% naive CPU implementation.

@@

\subsection{High Resolution Templates and FFT based Convolution}
\label{sec:fft} 
We generate high resolution templates with more than 250 HOG cells to capture
details of an object to give accurate 2D-3D matching. Since naive convolution takes
$O(nm)$ time where $n$ is the size of the signal and $m$ is the size of the template.
Whereas, FFT-based convolution takes $O\left( (n + m)\log (n+m) \right)$ time. 
For large $m$, high resolution template, we can gain computational advantage.

% We pad zero to the 2D image features by the largest kernel size. And we
% transform the kernel to match the padded image feature. To compute convolution,
% we compute element-wise product with the transformed kernel and sum all the
% features. Unlike \cite{Podlozhnyuk}, we do not need to shift the data nor
% kernel if we use simple mathematical trick. 
%   
% \begin{align}
%     x_{pad} & = \mathcal{P}_{n+m}(x)\\
%     y_{pad} & = \mathcal{P}_{n+m}(y)\\
%     x_{pad} \ast y_{pad} & = \mathcal{F}^{-1}(\mathcal{F}(x_{pad}) \circ \mathcal{F}(y_{pad}))
% \end{align}
% 
% Where $\ast$ is the circular convolution and $\circ$ is the Hadamard Product
% and $\mathcal{F}$ is the Fourier Transformation and $\mathcal{P}_n$ is the
% padding operation that append zeros to make vector of size $n$.
% 
% For each dimension of features, we transform kernel and image feature into
% frequency domain using Fast Fourier Transformation. We transform the image into
% frequency domain once and use the transformed image for all templates.
% 
% We implemented the convolution using FFT on GPU and got 10x speed up.

% [Figure for convolution time for different template resolution, CPU, GPU, naive convolution]
%------------------------------------------------------------------------

\section{Continuous Viewpoint Estimation using MCMC}
\label{sec:fine}
% Estimating viewpoint continuously to fine tune a initial matching
The NZ-WHO template matching method we have presented makes template generation
and evaluation computationally inexpensive. This means we can efficiently
explore the parameter space to find the best object pose, scale, model type and
camera focal length.

@@

Suppose that $x = \[p, m, f\]$ where $p$ is the 3D rotation of the CAD model and
$m$ is the CAD model index and $f$ is the focal length. Then we sample $x$ using 
% there are several advantages that was not possible such as \textbf{on-the-fly} template generation and making large number of templates  
\begin{figure}[t]
\centering
    \includegraphics[width=0.7\linewidth]{tuning2} \\ [-5pt]
    \caption{Different modes of Metropolis Hasting proposals, see Section \ref{sec:fine}}
 \label{fig:tuningmode}
\end{figure}
